---
title: 'NEPHU Surveillance Meetings: Enterics Updates'
author_details: "Alana Little, NEPHU (alana.little@austin.org.au)"
version_details: "v2.0, 21/09/2025"
output: 
  html_document: 
    code_folding: none
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = paste0("Enterics Updates ", Sys.Date()), output_dir = "../Output") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 1000px;
}

h1.title {
  font-weight: bold;
  font-size: 20px;
  color: #191d43;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #5bc788;
}

h2 {
  font-weight: bold;
  font-size: 18px;
  color: #5bc788;
}

h3 {
  font-weight: bold;
  font-size: 16px;
  color: #5bc788;
}

body {
  font-size: 14px;
}

.infobox {
  padding: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE)

options(knitr.kable.NA = '--')

here::i_am("Code/phar_checks_surveillance_meeting.Rmd")
```


```{r}
#
if (!require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse,
               lubridate,
               readxl,
               knitr,
               kableExtra,
               odbc,
               DBI,
               glue)
```


```{r}
# Epiweek for reporting
epiweek_start <- lubridate::ymd("2026-03-29")
epiweek_end   <- lubridate::ymd("2026-04-04")

# Date range for closed outbreaks
outbreak_start <- lubridate::ymd("2026-04-02")
outbreak_end   <- lubridate::ymd("2026-04-08")
```


```{r}
# Connect to PHAR and run queries
# Set useProxy = 0 when WFH and useProxy = 1 when in Austin office
con <- DBI::dbConnect(odbc::odbc(), "PHAR", useProxy = 0)

# New notifications during past epi week
all_enteric <- DBI::dbGetQuery(con,
                               glue::glue("SELECT * FROM dh_public_health.phess_release.caseevents_nrt
                                           WHERE (CONDITION_TYPE = 'Enteric Diseases' OR CONDITION = 'Rotavirus infection') 
                                           AND EVENT_CLASSIFICATION IN ('Confirmed', 'Probable')
                                           AND LPHU = 'NEPHU'
                                           AND EVENT_DATE >= DATE '{epiweek_start}'
                                           AND EVENT_DATE <= DATE '{epiweek_end}'"))

new_enteric <- all_enteric %>% 
  janitor::clean_names() %>% 
  #
  dplyr::filter(epi_ticked_for_follow_up_unit_last == "Local Public Health Unit") %>% 
  #
  dplyr::mutate(type = "New")

# Completed cases since last meeting
completed_enteric <- DBI::dbGetQuery(con,
                                     glue::glue("SELECT * FROM dh_public_health.phess_release.caseevents_nrt
                                                WHERE (CONDITION_TYPE = 'Enteric Diseases' OR CONDITION = 'Rotavirus infection') 
                                                AND EVENT_CLASSIFICATION IN ('Confirmed', 'Probable')
                                                AND LPHU = 'NEPHU'
                                                AND INVESTIGATION_COMPLETED_DATE >= DATE '{outbreak_start}'
                                                AND INVESTIGATION_COMPLETED_DATE <= DATE '{outbreak_end}'"))

completed_enteric <- completed_enteric %>% 
  janitor::clean_names() %>% 
  #
  dplyr::filter(epi_ticked_for_follow_up_unit_last == "Local Public Health Unit") %>% 
  #
  dplyr::mutate(type = "Completed")

# Closed outbreaks since last meeting
outbreaks_enteric <- DBI::dbGetQuery(con,
                                     glue::glue("SELECT * FROM dh_public_health.phess_release.outbreakevents_nrt
                                                WHERE CONDITION_TYPE = 'Enteric Diseases' 
                                                AND INVESTIGATION_STATUS = 'Completed'
                                                AND ASSIGNED_LPHU = 'North Eastern'"))

outbreaks_enteric <- outbreaks_enteric %>% 
  janitor::clean_names() %>% 
  #
  dplyr::filter(investigation_completed_date >= outbreak_start & investigation_completed_date <= outbreak_end) %>%
  dplyr::filter(!is.na(exposure_site)) %>% 
  #
  dplyr::select(event_id,
                outbreak_name,
                exposure_site,
                event_date,
                event_classification,
                condition,
                organism_cause,
                setting_where_exposure_occurred,
                investigation_completed_date) %>% 
  #
  dplyr::arrange(investigation_completed_date, outbreak_name)

# Close the connection to PHAR
DBI::dbDisconnect(con)

# Read in conditions reference list
conditions_reference <- readxl::read_xlsx(paste0(here::here(), "/Data/ConditionsReferenceList.xlsx"),
                                          sheet     = 1,
                                          guess_max = min(100000, Inf)) %>% 
  #
  janitor::clean_names() %>% 
  #
  dplyr::select(condition,
                chart_label)
```

## Closed Outbreaks Since Last Meeting

```{r}
#
outbreaks_enteric <- outbreaks_enteric %>% 
  dplyr::mutate(
    facility_name = stringr::str_sub(exposure_site, 32),
    #
    organism_cause = dplyr::case_when(
      organism_cause == "Viral (not laboratory confirmed)" ~ "Viral not confirmed",
      TRUE ~ organism_cause),
    #
    investigation_completed_date = format(investigation_completed_date, format = ("%d/%m/%Y"))) %>% 
  #
  dplyr::select(event_id,
                facility_name,
                setting_where_exposure_occurred,
                organism_cause,
                investigation_completed_date) %>% 
  #
  dplyr::arrange(investigation_completed_date, setting_where_exposure_occurred, organism_cause)
```

`r if(nrow(outbreaks_enteric) == 1) paste0("There has been ", nrow(outbreaks_enteric), "x closed outbreak since the last NEPHU weekly surveillance meeting on ", format(outbreak_start, format = "%d/%m/%Y"), ".")`

`r if(nrow(outbreaks_enteric) != 1) paste0("There have been ", nrow(outbreaks_enteric), "x closed outbreaks since the last NEPHU weekly surveillance meeting on ", format(outbreak_start, format = "%d/%m/%Y"), ".")`

**NEPHU enterics outbreaks closed between `r format(outbreak_start, format = "%d/%m/%Y")` and `r format(outbreak_end, format = "%d/%m/%Y")`**

```{r}
#
outbreaks_enteric %>% 
  knitr::kable(table.attr = "style = \"color: black;\"",
               align      = "lllll",
               col.names  = c("PHESS ID",
                              "Facility name",
                              "Facility type",
                              "Organism",
                              "Closure date")) %>% 
  #
  kableExtra::kable_styling(full_width = TRUE,
                            position   = "center",
                            html_font  = "Arial",
                            font_size  = 12) %>% 
  #
  kableExtra::row_spec(row       = nrow(outbreaks_enteric),
                       extra_css = "border-bottom: 2px solid lightgray;")
```

***

## New and Completed Cases in Past Epiweek

```{r}
# Select cases for review
review_enteric <- new_enteric %>% 
  dplyr::bind_rows(completed_enteric) %>% 
  #
  dplyr::left_join(conditions_reference, by = "condition") %>% 
  #
  dplyr::mutate(
    condition_name = dplyr::case_when(
      condition == "Food-borne or water-borne illness" ~ "Potential FBWB Illness",
      TRUE ~ chart_label),
    #
    sex = stringr::str_sub(sex, 1, 1),
    #
    age_sex = paste0(age_years, sex),
    #
    suburb = stringr::str_to_title(city),
    #
    lga = stringr::str_sub(lga, end = -4),
    #
    notes = "") %>%
  #
  dplyr::arrange(condition, event_date) %>%
  #
  dplyr::distinct(event_id, .keep_all = TRUE) %>% 
  #
  dplyr::select(condition_name,
                type,
                event_id,
                event_classification,
                age_sex,
                suburb,
                lga,
                notes)
```

There are `r review_enteric %>% dplyr::filter(type == "New") %>% nrow()`x newly signed out enterics cases from the past epiweek and there have been `r review_enteric %>% dplyr::filter(type == "Completed") %>% nrow()`x enterics cases from previous epiweeks completed since the last NEPHU weekly surveillance meeting.

**NEPHU newly signed out cases between `r format(epiweek_start, format = "%d/%m/%Y")` and `r format(epiweek_end, format = "%d/%m/%Y")` and completed enterics cases between `r format(outbreak_start, format = "%d/%m/%Y")` and `r format(outbreak_end, format = "%d/%m/%Y")`**

```{r}
#
review_enteric %>% 
  knitr::kable(table.attr = "style = \"color: black;\"",
               align      = "llllllll",
               col.names  = c("Condition",
                              "Type",
                              "PHESS ID",
                              "Classification",
                              "Age/Sex",
                              "Suburb",
                              "LGA",
                              "Notes")) %>% 
  #
  kableExtra::kable_styling(full_width = TRUE,
                            position   = "center",
                            html_font  = "Arial",
                            font_size  = 12) %>% 
  #
  kableExtra::row_spec(row       = nrow(review_enteric),
                       extra_css = "border-bottom: 2px solid lightgray;")
```

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
